**Sequencing 16S rRNA gene fragments using the PacBio SMRT DNA sequencing system**

### 2016-02-05 Schloss_PacBio16S_PeerJ_2016.Rmd
Edit Abstract section

```{r knitr_settings, eval=TRUE, echo=FALSE, cache=FALSE}
opts_chunk$set("tidy" = TRUE)
opts_chunk$set("echo" = FALSE)
opts_chunk$set("eval" = TRUE)
opts_chunk$set("warning" = FALSE)
opts_chunk$set("cache" = FALSE)

@@

clrs <- rainbow(length(regions))
names(clrs) <- regions

library(rjson)
```

@@

## Abstract [ Revise ]
Over the past 10 years, microbial ecologists have largely abandoned sequencing
16S rRNA genes by the Sanger sequencing method and have instead adopted highly
parallelized sequencing platforms. These new platforms, such as 454 and
Illumina's MiSeq, have allowed researchers to obtain millions of high quality,
but short sequences. These platforms have allowed researchers to significantly
improve the design of their experiments. The tradeoff has been the decline in
the number of full-length reference sequences that are deposited into databases.
To overcome this problem, we tested the ability of the PacBio Single Molecule,
Real-Time (SMRT) DNA sequencing platform to generate sequence reads from the 16S
rRNA gene. We generated sequencing data from the V4, V3-V5, V1-V3, V1-V6, and
V1-V9 variable regions from within the 16S rRNA gene from a synthetic mock
community and natural samples collected from human feces, mouse feces, and soil.
The synthetic mock community allowed us to assess the actual sequencing error
rate and how that error rate changed when different curation methods were
applied. We developed a simple method based on sequence characteristics and
quality scores to reduce the observed error rate for the V1-V9 region from 2.16%
to 0.32%. Unfortunately, this error rate was still 16-times higher than the
error rate that has been observed for the shorter reads generated by 454 and
Illumina's MiSeq sequencing platforms. Although the longer reads frequently
provided better classification, the wider adoption of this approach for 16S rRNA
gene sequencing is likely limited by its high sequencing error and low yield of
sequencing data relative to the other available platforms.

@@

```{r error_profile}
error_profile <- fromJSON(scan("data/process/error_profile.json", quiet=T, what="character"))
```

@@

```{r basic}
error_summary <- read.table(file="data/process/error_summary.tsv", header=T)

raw_to_basic_lost <- 100 * (1 - error_summary[error_summary$method == "basic","nseqs"]/error_summary[error_summary$method == "raw","nseqs"])
```

@@

```{r}
sobs_data <- read.table(file="data/process/sobs_table.tsv", header=T)
perfect_otu <- 19
```

@@

```{r, classificationSummary}
taxonomy_data <- read.table(file="data/process/taxonomy_depth_analysis.tsv", header=T)
```

@@

```{r}
one_offs <- read.table(file="data/process/non_random_analysis.tsv", header=T)
percent_one_off <- 100 * prop.table(as.matrix(one_offs), 2)

>>>>

```{r knitr_settings, eval=TRUE, echo=FALSE, cache=FALSE}
library(rjson)

opts_chunk$set("tidy" = TRUE)
opts_chunk$set("echo" = FALSE)
opts_chunk$set("eval" = TRUE)
opts_chunk$set("warning" = FALSE)
opts_chunk$set("cache" = FALSE)

@@

error_profile <- fromJSON(scan("data/process/error_profile.json", quiet=T, what="character"))
error_summary <- read.table(file="data/process/error_summary.tsv", header=T)
sobs_data <- read.table(file="data/process/sobs_table.tsv", header=T)
taxonomy_data <- read.table(file="data/process/taxonomy_depth_analysis.tsv", header=T)
one_offs <- read.table(file="data/process/non_random_analysis.tsv", header=T)
```

@@

## Abstract
Over the past 10 years, microbial ecologists have largely abandoned sequencing
16S rRNA genes by the Sanger sequencing method and have instead adopted highly
parallelized sequencing platforms. These new platforms, such as 454 and
Illumina's MiSeq, have allowed researchers to obtain millions of high quality,
but short sequences. The result of the added sequencing depth has been significant
improvements in experimental design. The tradeoff has been the decline in
the number of full-length reference sequences that are deposited into databases.
To overcome this problem, we tested the ability of the PacBio Single Molecule,
Real-Time (SMRT) DNA sequencing platform to generate sequence reads from the 16S
rRNA gene. We generated sequencing data from the V4, V3-V5, V1-V3, V1-V5, V1-V6,
and V1-V9 variable regions from within the 16S rRNA gene using DNA from a synthetic mock community and natural samples collected from human feces, mouse feces, and soil. The mock community allowed us to assess the actual sequencing error rate and how that error rate changed when different curation methods were
applied. We developed a simple method based on sequence characteristics and
quality scores to reduce the observed error rate for the V1-V9 region from
`r format(error_summary[error_summary$region == "V1V9" & error_summary$method == "raw", "error"], digits=2, nsmall=2)` to `r format(error_summary[error_summary$region == "V1V9" & error_summary$method == "precluster", "error"], digits=2, nsmall=2)`%. This error rate is comparable
to what has been observed for the shorter reads generated by 454 and
Illumina's MiSeq sequencing platforms. Although the per base sequencing cost is
still significantly more than that of MiSeq, the prospect of supplementing reference databases with full-length sequences from organisms below the limit of detection from the Sanger approach is exciting

@@

@@

```{r basic}
raw_to_basic_lost <- 100 * (1 - error_summary[error_summary$method == "basic","nseqs"]/error_summary[error_summary$method == "raw","nseqs"])
```

@@

```{r}
perfect_otu <- 19
```

@@

@@

```{r}
percent_one_off <- 100 * prop.table(as.matrix(one_offs), 2)
